---
title: "Lab 1"
author: "Alice"
date: "1/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Attach Packages

```{r}
library(tidyverse)
library(sf)
library(leaflet)
library(tmap)
```
2. Review: Reading, exploring, wrangling and visualizing data

a. Read in the Data

```{r}

wine <- read_csv("wine_data.csv")

```

b. wrangling
```{r}

#Select chooses columns. Can use -select to exclude columns instead. Columns go in the order you write them down after select
#one equal sign changes name, two equal signs looks for match. Overwrite province to state
#vertical line means OR. Double == means 'matches'
#Mutate adds column
#Arrange sorts in decreasing points per dollar

wine_new <- wine %>% 
  select(country, province, winery, region_1, points, price) %>%  
  rename(state = province) %>% 
  filter(state == "California"|state == "Oregon"|state == "Washington") %>%  
  mutate(ppd = points/price) %>% 
  arrange(-ppd)


  
```

c. Use groub_by() and summarize() to find summary statistics by group
```{r}
#Grouping data in R's brain and then finding summary statistics. We won't see these groups, but R will be doing it behind the scenes. It will calculate things based on different levels. 
#R will recognize each unique state and treat them each as seperate groups for following calculations
#create a data table with mean values of price per dollar based on different group.
#na.rm = TRUE means remove any NA values that exist

wine_summary <- wine_new %>% 
  group_by(state) %>% 
  summarize(
    mean_ppd = mean(ppd, na.rm = TRUE),
    max_ppd = max(ppd, na.rm = TRUE)
  )
  
```

d. A few gg-graphics

```{r}
#Naming things is not necessary but is good practice
#When we use ggplot we have to tell it 3 things. 1. Tell it we're using GGplot 2. Tell it what data we're using. 3. Tell it what type of plot
#Histogram doesn't require y
#piping adds additional functions that operate in function
#within ggplot we use plus signs to show we're adding additional levels
#For histograms, we use fill to distinguish colors
#If we want to do something based on a variable, we have to use aes
#see that maybe having all wineries seperately doesn't make sense 
#so we run by state (facet_wrap)
#By default the scales are the same
#We can override that by saying scales = "free"


wine_hists <- ggplot(wine_new, aes(x=ppd))+
  geom_histogram(aes(fill=state))+
  facet_wrap(~state, scales = "free")

#call it
wine_hists


```

new graph: scatterplot graph of price vs. points, with color determined by state, adding a linear model trendline on top of it

```{r}

#geom_point gives scatterplot
#POINTS USE COLOR, BARS USE BILL, AND COLOR IS THE LINE COLOR
#If not referencing a variable, put things OUTSIDE of aes
#Alpha is for transparency
#You can overlay models over the graph--> geom_smooth
#If you want to choose colors manually use scale_color_manual
#You could also do a gradient
#c stands for combination
#again, facetwrap is to seperate into seperate graphs each by state
#we can throw some themes in there

wine_points <- ggplot(wine_new, aes(x=price, y=points)) +
  geom_point(aes(color = state), alpha = 0.5) +
  geom_smooth(color = "black", method = "lm", se = FALSE) +
  scale_color_manual(values = c("royalblue", "mediumorchid1", "orange")) +
  facet_wrap(~state) +
  theme_light()

wine_points

```

2. California Counties (spatial data) using sf

```{r}

#st because sf functions will eventually become spatiotemporal functions
#"." tells it we're already there. It's already in our working directory
#layer tells it to call all these similarly named things


ca_counties <- st_read(dsn = ".", layer = "CA_Counties_TIGER2016" )

ca_land <- ca_counties %>% 
  select(NAME, ALAND)

#call in income database
#want to merge these databases through Join function, but first, want to check that the variables are named in ways that match
#Not just headings but also categories

ca_pop_inc <- read_csv("ca_pop_inc.csv") %>% 
  rename(NAME=COUNTY)

#full join is your safest bet because it keeps every row even if it doesn't have a match in the other dataframe
ca_df <- full_join(ca_land, ca_pop_inc) %>% 
  select(NAME, MedFamilyIncome) #remember geometries are sticky


#plot(ca_counties)
#above is commented out because it takes a long time to load

#sf uses sticky geometries. Means you don't have to worry about dealing with the geometry. It just stays there
```

d. Use geom_sf to make a map in ggplot

```{r}

ca_income <- ggplot(ca_df) +
  geom_sf(aes(fill = MedFamilyIncome))

ca_income

#we need data to match the coordinate system

ca_df_transform <- st_transform(ca_df, crs = 4326)

leaflet(ca_df_transform) %>% 
  addTiles() %>% 
  addPolygons()

#Creates INTERACTIVE MAP!

```

